Global Climate of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and Spitzer Phase Curve Observations

### 2017-10-25 ms.tex
DO NOT FORGET TO PUSH YOUR CHANGDS

\section{Light Curve Fits}
We fit a two-component model to the light curves. One component models the astrophysical signal (the planet's thermal phase variation and transit), and the other component models the systematic noise introduced by time-dependent changes in instrument performance. For each light curve, we fit the physical and systematic components simultaneously, such that the total observed flux as a function of time is given by $F(t) = F_\mathrm{physical}(t) \times F_\mathrm{sys}(t)$. For the \HST\ data, where we observed two phase curves and two additional eclipses, we constrain the physical parameters to be the same for all visits, but allow some of the systematics parameters to vary (for more details see \S\,\ref{sec:hst_sys}). We fit the WFC3 band-integrated (``white" light curve"), as well as spectroscopic light curves created from 10 wavelength bins uniformly spaced at $0.05 \mu$m intervals.

\subsection{Astrophysical Signal}
We assume the astrophysical signal $F_\mathrm{physical}(t)$ has the following form:
\begin{equation}
	F_\mathrm{physical}(t) = F_s (1 + \alpha) \times T(t) + F_p(t) 
\end{equation}
where $F_s$ is the stellar flux, $T(t)$ is the transit model (the fraction of the stellar disk that is visible), $\alpha$ is additional fractional flux from the companion star, and $F_p(t)$ is the planet flux. FIXME what is alhpha. The flux from the planet is given by
\begin{equation}
F_p(t) = \pi I(t)[R(t)]^2
\end{equation}
where $I(t)$ is the disk-integrated planet intensity and $R(t)$ is the planet's effective radius (which changes with orbital phase due to tidal distortion).

We calculate the transit model with the \texttt{batman} package \citep{kreidberg15a}. Many of the physical parameters are tightly constrained by \cite{southworth15}, so we fixed the orbital period, time of inferior conjunction, orbital inclination, and ratio of semi-major axis to stellar radius to the previously published values ($P = 0.925545613$ day, $t_0 = 2456836.2964455 \mathrm{BJD_{TDB}}$, $i = 87.3^\circ$, and $a/R_s = 2.999$). As a test, we fit for these parameters with the Spitzer Channel 2 light curve, which has the best phase coverage and least systematic noise of the three data sets. We found that the transit parameters are consistent with the \cite{southworth15} results, so we proceeded with the remainder of the analysis holding those parameters fixed.  The free parameters for the transit model were the transit depth $r_p$ and a linear limb darkening parameter $u$.

We explored several different models for the planet's thermal phase variation $P(t)$, using climate maps from the open-source Python package \texttt{SPIDERMAN}. This package allows users to input a climate map (temperature or brightness as a function of latitude and longitude), and generate the corresponding thermal phase curve over a specified wavelength range. We tested three different maps:  a two-temperature map, with a uniform dayside temperature $T_d$ and a uniform nightside temperature $T_n$; a map generated with spherical harmonics; and the physically-motivated kinematic model from \cite{zhang16}, which has just three free parameters (the nightside temperature $T_n$, the change in temperature from day-to-night side $\Delta_T$, and the ratio of radiative to advective timescales $\xi$). These models each have different strengths. The kinematic model incorporates the most prior knowledge of physics, and even though it makes several simplifying assumptions (e.g, that the wind speed and direction are constant), it closely reproduces GCM simulations over a wide parameter space \citep{zhang16}.   The spherical harmonics map is the most flexible and makes the fewest assumptions about the true climate. The two-temperature model is the simplest and has the fewest free parameters.  In addition to these maps, we also fit the phase variation with a sinusoid, which has been used for many past analyses of exoplanet phase curves (FIXME).
%We explored adding an additional sine term to model the thermal phase variation, but the additional degrees of freedom are not justified according to the Bayesian Information Criterion (BIC).

In addition to the changing temperature of the planet's surface, the visible surface area is also expected to change as a function of orbital phase.  WASP-103b is so close to its host star that it is likely tidally distorted and partially filling its Roche lobe \citep{gillon14}, which introduces ellipsoidal variations $E(t)$ in the light curve.  Our light curves are not precise enough to detect the ellipsoidal variability, but we include a model prediction $E(t)$ to avoid introducing any bias in the thermal phase variation. We model $E(t)$ using the analytic expression from \cite{leconte11}, which

FIXME doppler beaming?

\begin{figure}
\includegraphics[width = 0.5\textwidth]{Figures/ellipsoidal.pdf}
\caption{Projected area of the planet as a function of orbital phase, normalized to unity at phase zero. The area variation was predicted analytically using the model from \citep{leconte11}.}
\label{fig:ellipsoidal}
\end{figure}

>>>>

\section{Light Curve Fits}
We fit a two-component model to the light curves. One component models the astrophysical signal (the planet's thermal phase variation and transit), and the other component models the systematic noise introduced by time-dependent changes in instrument performance. For each light curve, we fit the physical and systematic components simultaneously, such that the total observed flux as a function of time is given by $F(t) = F_\mathrm{physical}(t) \times F_\mathrm{sys}(t)$. For the \HST\ data, where we observed two phase curves and two additional eclipses, we constrain the physical parameters to be the same for all visits, but allow some of the systematics parameters to vary (for more details see \S\,\ref{sec:hst_sys}). We fit the WFC3 band-integrated (``white" light curve"), as well as spectroscopic light curves created from 10 wavelength bins uniformly spaced at $0.05\,\mu$m intervals between $1.15$ and $1.65\,\mu$m.

\subsection{Astrophysical Signal}
We assume the measured astrophysical signal $F_\mathrm{physical}$ has the following form:
\begin{equation}
	F_\mathrm{physical}(t, \lambda) \propto F_s(\lambda) (1 + \alpha(\lambda)) \times T(t, \lambda) + F_p(t, \lambda) 
\end{equation}
where $F_s(\lambda)$ is the stellar flux as a function of wavelength, $\alpha(\lambda)$ is the additional fractional flux from the companion star, $T(t, \lambda)$ is the transit model (the fraction of the stellar disk that is visible at time $t$), and $F_p(t, \lambda)$ is the disk-integrated planet flux. These terms are described in more detail below. We assume the stellar flux $F_s(\lambda)$ is a $6110\,\mathrm{K}$ blackbody \citep{gillon14}. The flux from the planet is given by
\begin{equation}
F_p(t) = I(t, \lambda)[R(t)]^2
\end{equation}
where $I$ is the disk-integrated planet intensity and $R$ is the planet's effective area (which changes with orbital phase due to tidal distortion).  In our final fit, we divide by the total stellar flux (primary and background stars), such that the astrophysical signal is:
\begin{equation}
	F_\mathrm{physical}(t) = T(t) + F_p(t, \lambda) /[F_s(\lambda) (1 + \alpha (\lambda))]
\end{equation}

We calculate the transit model $T(t)$ with the \texttt{batman} package \citep{kreidberg15a}. Many of the physical parameters are tightly constrained by \cite{southworth15}, so we fixed the orbital period, time of inferior conjunction, orbital inclination, and ratio of semi-major axis to stellar radius to the previously published values ($P = 0.925545613$ day, $t_0 = 2456836.2964455 \mathrm{BJD_{TDB}}$, $i = 87.3^\circ$, and $a/R_s = 2.999$). As a test, we fit for these parameters with the Spitzer Channel 2 light curve, which has the best phase coverage and least systematic noise of the three data sets. We found that the transit parameters are consistent with the \cite{southworth15} results, so we proceeded with the remainder of the analysis holding those parameters fixed.  The free parameters for the transit model were the transit depth $r_p$ and a linear limb darkening parameter $u$.

We explored several different models for the planet's disk-integrated intensity $I(t, \lambda)$, using climate maps from the open-source Python package \texttt{SPIDERMAN} \citep{louden17}. This package allows users to input a climate map (temperature or brightness as a function of latitude and longitude), and generate the corresponding planet-to-star flux for an observation at time $t$, over a specified wavelength range. We tested three different maps: a two-temperature map, with a uniform dayside temperature $T_d$ and a uniform nightside temperature $T_n$; a map generated with spherical harmonics; and the physically-motivated kinematic model from \cite{zhang16}, which has just three free parameters (the nightside temperature $T_n$, the change in temperature from day-to-night side $\Delta_T$, and the ratio of radiative to advective timescales $\xi$). These models each have different strengths. The kinematic model incorporates the most prior knowledge of physics, and even though it makes several simplifying assumptions (e.g, that the wind speed and direction are constant), it closely reproduces GCM simulations over a wide parameter space \citep{zhang16}.   The spherical harmonics map is the most flexible and makes the fewest assumptions about the true climate. The two-temperature model is the simplest and easiest to compare to one-dimensional models of the dayside, which assume a uniform temperature/pressure profile.  In all cases, we assumed that the planet is tidally locked, such that each orbital revolution corresponds to one complete rotation on its spin axis. %In addition to these maps, we also fit the phase variation with a sinusoid, which has been used for many past analyses of exoplanet phase curves (FIXME).
%We explored adding an additional sine term to model the thermal phase variation, but the additional degrees of freedom are not justified according to the Bayesian Information Criterion (BIC).

In addition to temperature, the planet's visible surface area $[R(t)]^2$ is also expected to vary with orbital phase due to tidal distortion \citep{gillon14}. In the observed light curve, changes in planet radius are degenerate with changes in temperature. To address the possible effects of radius variability on our data, we predict the variation using the analytic formula from \cite{leconte11b}, equation B.9, which computes the projected area of a triaxial ellipsoid. We assume a Love number $\kappa_n = 0.653$ based on Table B.5 of \cite{leconte11a}, for a highly irradiated, $1.5 R_\mathrm{J}$ planet aged 5 Gyr. The predicted ellipsoidal variability is shown in Figure\,\ref{fig:ellipsoidal}. At quadrature, the project radius is $8\%$ larger than at phase zero (mid-transit). FIXME: say this doesn't matter.

FIXME: say what the dilution factor is.
%Our light curves are not precise enough to detect the ellipsoidal variability, but we include a model prediction $E(t)$ to avoid introducing any bias in the thermal phase variation. We model $E(t)$ using the analytic expression from \cite{leconte11}, which 
%FIXME doppler beaming?

\begin{figure}
\includegraphics[width = 0.5\textwidth]{Figures/ellipsoidal.pdf}
\caption{Projected area of the planet as a function of orbital phase, normalized to unity at phase zero. The area variation was predicted analytically using the model from \citep{leconte11b}.}
\label{fig:ellipsoidal}
\end{figure}
